Algebraic Groups\vspace{-1ex}

### 2019-10-07 Algebraic Geometry/Algebraic-Geometry.tex
update

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}
\let\O\relax
\newcommand*{\O}{\calO}
\newcommand{\A}[1]{\bbA^{#1}}
\let\P\relax
\newcommand{\P}[1]{\bbP^{#1}}
\newcommand{\eqdef}{\stackrel{\text{def}}{=}}

@@

	Thus $J'=(1)$. So we can write $1=\frac{f}{g^N}$. Thus $g^k(f-g^N)=0$ in $k[x_i]$ and since $g$ isn't nilpotent, $f=g^N$.
\end{prf}
\begin{cor}
	There is a lattice anti-isomorphism between the radiacal ideals in $k[x_i]$ and Zariski-closed 
	subsets $Z\subseteq k^n$ via the maps $J\to Z(J)$ and $Z\mapsto I(Z)$.
\end{cor}
\begin{cor}
	For any ideal $J\subseteq k[x_i]$, 
	\[\sqrt{J}=\bigcap_{\text{maximal } \frakm\supset J}\frakm\]
\end{cor}
\begin{rmk}
	``The functions that vanish at the zero locus of $J$ are precisely those that vanish at all the points of $J$''.
\end{rmk}
\begin{cor}
	$D(g)\subseteq k^n$. Then the map 
	\[k[x_i][1/g]\to \Hom(D(g),k)\]
	via the map 
	\[\frac{f}{g^N}\mapsto \left(x\mapsto \frac{f(x)}{g(x)^n}\right)\]
	is injective.
\end{cor}

@@

>>>>

% Document-Specific Macros

@@

	Thus $J'=(1)$. So we can write $1=\frac{f}{g^N}$. Thus $g^k(f-g^N)=0$ in $k[x_i]$ and since $g$ isn't nilpotent, $f=g^N$.
\end{prf}
\begin{cor}
	There is a lattice anti-isomorphism between the radiacal ideals in $k[x_i]$ and Zariski-closed 
	subsets $Z\subseteq k^n$ via the maps $J\to Z(J)$ and $Z\mapsto I(Z)$.
\end{cor}
\begin{cor}
	For any ideal $J\subseteq k[x_i]$, 
	\[\sqrt{J}=\bigcap_{\text{maximal } \frakm\supset J}\frakm\]
\end{cor}
\begin{rmk}
	``The functions that vanish at the zero locus of $J$ are precisely those that vanish at all the points of $J$''.
\end{rmk}
\begin{cor}
	$D(g)\subseteq k^n$. Then the map 
	\[k[x_i][1/g]\to \Hom(D(g),k)\]
	via the map 
	\[\frac{f}{g^N}\mapsto \left(x\mapsto \frac{f(x)}{g(x)^N}\right)\]
	is injective.
\end{cor}

@@

\section{October 7th, 2019}
\subsection{Questions/Discussions}
\subsubsection{Initial and final objects}
We asked before whether the space with functions $(X,\O^{\text{loc. constant}})$ is an initial or terminal 
object. Adam asserts that it is a final object in the category of spaces with functions where the underlying space is $X$ (I believe this 
is sheaves over $X$).

Notice we can't use all continus maps where $k$ has the discrete topology, since $0\in k$ is not closed.

\subsubsection{Subalgebras and rings of functions}
Assume that $f:B\to A$. This gives us a nice map $\tilde f:\Spec A\to \Spec B$. One question we may have is 
``if $f$ is injective, does this imply that $\tilde f$ is surjective?''

Consider an example: Say $B=k[t]\hookrightarrow k[s,t]=A$. Then any function on $A$ looks like $(s-a,t-b)$
and the map induced on functions is just projection, so this gives us the map $(t-b)$, which is all the maps on $B$.

Another example: consider the map $\bbC[t]\to\bbC[s]$ sending $t\to s^2$. Then this induces a map $z\to z^2$ from $\bbC\to \bbC$ (why?)
which is again surjective.

Next consider $k[x,u]/(xy-1)=k[x][1/x]=A$, which is a hyperbola over $\bbR$. Then the localization map $B=k[x]\hookrightarrow A$
induces a map that is basically the identity everywhere \textit{except zero}. So it is \textbf{not surjective.}

\brk

Some properties to notice: examples one and two are \textit{flat} extensions. The second is a \textbf{finite} extension. The 
third is neither. We will investigate what is going on further later on.

One idea: consider whether the map $X=\Spec B\to \A 2\setminus(\A1\setminus\{0\})$ exists. One of the things that he keeps questioning is 
whether the target spaces is open or closed as a subset of affine space. (Note that a map is proper if it sends closed to closed).

\subsection{Back to affine varieties}
Continuing our proof/discussion from last time, we were considering $\varphi:X\to\A n$ which we said was a morphism iff each coordinate $\phi_i:X\to k$ is regualr.
(This basically follow since $\varphi$ is continuous and sends regular functions to regular functions.

For the regularity, consider $U\subseteq\A n$. Then $U$ admits a cover of $D(g)$, so it suffices to check where $\phi$
pulls back regular maps on $D(g)$. Of course this is $k[x_1,\dots,x_n][1/g]$! So consider the image of 
\[\frac{1}{g^N}\sum a_ix^i\]
and its image in $\phi^{-1}(D(g))=D(\phi^\ast(g))$ is 
\[\frac{\sum a_i(\phi^\ast x)^i}{(\phi^\ast g)^N}\]

\begin{rmk}
	Big idea: We started with the ``dream'': that there is a correspondence between the algebra and the geometry. This is 
	our main guiding principle, so we know we've found the ``right'' topology when we have found one that supports this dream.
	This is an answer to the question ``why is the Zariski topology not just a degenerate case?''
\end{rmk}

\subsubsection{Affine varieties in general}
Let $J\subseteq k[x_1,\dots,x_n]$ be a radical ideal. We know this corresponds (uniquely!) 
to a subset $Z\subseteq \A n$. Here we can consider $(Z,\O_Z)$ where $Z$ is a subspace of $\A n$.

Then take any closed $W\subseteq Z$, which is the intersection (by definition of the topology)
\[\cap_{i\in I}Z(f_i)\quad f_i\in h[x_i]/J\]
and then $\O_Z(D(g))=\frac{k[x_i]}{J}[1/g]$ (note we used the nullstellensatz here!).

So then we claim that $Z$ is an affine variety. To see this, consider a map $\varphi:X\to Z$ and the composition
\[X\xrightarrow{\varphi} Z\hookrightarrow \A n\]
so topologically $X$ cfactors through $Z$ if and only if $J\subseteq \O^{\A n}(\A n)$ maps to zero in $\O_X(X)$.

The takeaway here is that a morphism $X\to\A n$ factors through $Z$ topologically if and only if it factors in the categorification of 
spaces with functions.

### 2019-10-07 Algebraic Groups/Algebraic-Groups.tex
update

% Document-Specific Macros
\DeclareMathOperator{\1}{\mathbbm{1}}
\DeclareMathOperator{\GL}{GL}
\DeclareMathOperator{\Sym}{Sym}
\DeclareMathOperator{\Spec}{Spec}
\newcommand{\red}[1]{{#1}_{\text{red.}}}

@@

>>>>

% Document-Specific Macros

@@

\section{October 7, 2019}
Today we will be primarly concerned with 
\subsection{Group actions}
Let $G$ be an algebraic group over $k$. 
\begin{defn}
	A \textbf{group action} of $G$ on a scheme $X$ over $k$ is the data of a morphism
	\[\mu:G\times X\to X\]
	such that the usual axioms hold. That is,
	\begin{center}
		\begin{tikzcd}
			G\times G\times X\ar[r,"m\times\id"]\ar[d,"\id\times \mu"] & G\times X\ar[d,"\mu"]\\
			G\times X\ar[r,"\mu"] & X
		\end{tikzcd}
		\begin{tikzcd}
			\Spec k\times x\ar[r,"e\times\id"]\ar[rd,"\sim"] & G\times X\ar[d,"\mu"]\\
			& X
		\end{tikzcd}
	\end{center}
\end{defn}
\begin{rmk}
	Apparently it was an exercise already to show that this is equivalent to an action of $h_G$ on $h_X$.
\end{rmk}
\begin{rmk}
	The map $(g,x)\mapsto(g,gx)$ is an automorphism of $G\times X$, so if $p_2:G\times X\to X$ is projection,
	\begin{center}
		\begin{tikzcd}
			G\times X\ar[rr,"\sim"] \ar[dr,"\mu"] & & G\times X\ar[dl,"p_2"]\\
			& X &
		\end{tikzcd}
	\end{center}
	commutes.
\end{rmk}
\begin{defn}
	Let $X$ and $Y$ be schemes over $k$ with a $G$ action. Then a \textbf{$G$-equivariant morphism} $f:X\to Y$ 
	is one such that for all $g\in G$, the following commutes:
	\begin{center}
		\begin{tikzcd}
			G\times X\ar[r,"\id\times f"]\ar[d,"\mu_X"] &G\times Y \ar[d,"\mu_Y"]\\
			X\ar[r,"f"] & Y
		\end{tikzcd}
	\end{center}
\end{defn}
\subsubsection{Some examples}
\begin{itemize}
	\item $G$ actions on itself by multiplication or conjugation.
	\item $\Gm$ acts on $\A1$. Geometrically, we are just looking at $k^\ast$ acting on $k$ by scaling.
	Algebraically, we want a map $\mu\Gm\times\A1\to\A1$ given by the map of algebras:
	\[k[x]\xrightarrow{\mu^\ast}k[t]_t\otimes k[x]\quad\text{via}\quad x\mapsto tx\]
	Functorially, if $S$ is a scheme over $k$, then $\Gm(S)=\Gamma(S,\O_S)^\ast$ which acts on $\A1(S)=\Gamma(S,\O_S)$,
	again by scaling.
	\item You can consider $\GL_n$ action on $\A n$ by multiplicaiton or on $\A{n\times n}$ via multiplication or conjugation.
\end{itemize}
\subsubsection{Orbits and Stabilizers}
Let $G$ be an algebraic group over $k$ action on a scheme $X$ over $k$. Let $x\in X(k)$. Then we have a map
\[\mu_x:G\times \Spec k\xrightarrow{\id\times x}G\times X\xrightarrow{\mu}X\]
where 
\[g\mapsto (g,x)\mapsto gx.\]
\begin{defn}
	The \textbf{orbit of $x$} is $Gx=\mu_x(G)\subseteq |X|$ set-theoretically. The \textbf{stabilizer of $x$ in $G$} 
	is $G_x=\mu_x^{-1}(x)\subseteq G$. 
\end{defn}
\begin{rmk}
	$G_x$ is always a closed algebraic subgroup of $G$.
\end{rmk}
\begin{prop}
	$\mu_x(G)$ is open in its closure in $|X|$.
\end{prop}
Recall first the following:
\begin{thm}[Chevalley's Theorem (different?)]
	If $f:X\to Y$ is a map of schemes of finite type over $k$, then $f(X)\subseteq Y$ is contstructible (i.e. 
	is a disjoint union of finitely many locally closed subsets).

	\noindent\textit{Recall that locally closed means closed in an open subspace.}
\end{thm}
\begin{cor}
	\textit{Maybe a definition:} The orbit $Gx\subseteq \operatorname{im}(\mu_x)\subseteq X$. If 
	$G$ is reduced, then $Gx$ is reduced.
\end{cor}

\subsubsection{Applications}
Say that $\ch k=p$. Then $\mu_p$ acts on $\Gm$ by multiplication.

\brk

$\Gm$ acts on $\A1$ with two orbits: $\Gm\cdot 1=\{x\ne 0\}$ and $\Gm\cdot 0=\{0\}$. The stabilizers are 
$G_1=1$ and $G_0\cong\Gm$.

\brk

Consider $G$ acting on $\A2$ via $t(x,y)=(tx,t^{-1}y)$. Then the orbits are hyperbolas! There is also a notion of closed 
orbits that I didn't quite catch. Also apparently the orbit-stabilizer statement is easy to see in geometry via 
a fiber bundle $G$ over $Gx$ where the fiber over $x$ is $G_x$.

\begin{prop}
	If $\phi:G\to H$ is a homomorphism of algebraic groups, then $\phi(G)\subseteq |H|$ is closed.
\end{prop}
\begin{rmk}
	The proof included reducing first to $k=\bar k$. The trick here is to consider the group action induced by $\phi$
	and then consider the map $\mu_{e_H}$ of this action. Then $\mu_{e_H}(G)=\phi(G)$ and one can prove that this is closed.
\end{rmk}

In particular, we have that \textbf{subgroups of an algebraic group are always closed.} Note that this stands in stark contrast 
to Lie theory where you get
